---
title: "ST451 Computer Workshop with code for the activity - Week 5"
output: pdf_document
---

### RStan

#### Demonstration

```{r}
library("rstan")
rstan_options(auto_write = TRUE)
```

First simulate data and put them in a list.

```{r}
set.seed(10)
N=1000;
mu = 5;
sigma = 4;
obs = rnorm(N,mu,sigma);
mean(obs)
sd(obs)

toy_dat <- list(N=N, y=obs)
```

For the model see the file `ToyExample.stan`. 

Next we run the automatic Variational Bayes algorithm

```{r}
ToyModel <- stan_model(file='ToyExample.stan')
vb(ToyModel,data=toy_dat)
```


#### Activity

The dataset contains cases from a study that was conducted between, 1958 and 1970 at the University of Chicago's Billings Hospital on the survival of patients who had undergone surgery for breastcancer.

Four variables:
  * Age of patient at time of operation 
  * Patient's year of operation (year - 1900)
  * Number of positive axillary nodes detected 
  * Survival status (1 = the patient survived 5 years or longer, 2 = the patient died within 5 year)

First we load the data and fit a logistic regression model with MLE

```{r}
library(data.table)
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/haberman/haberman.data"
biopsy <- fread(url, sep = ",", header = FALSE)
column.names <- c('age', 'year', 'nodes', 'survival')
setnames(biopsy, column.names)
biopsy$survival[biopsy$survival==2] = 0;
summary(biopsy)
fit1 = glm(survival~nodes,family=binomial,data=biopsy)
summary(fit1)
```

Normal priors can be assigned on the $\beta$ coefficients with zero mean and some reasonably large variance -say $100$.

The task is to obtain the mean field Variational Bayes approximation to the posterior of $\beta$ using ADVI in RStan.  

The model can be written as

$$y_i\sim Bernoulli(\pi_i),\;\; i=1,...,N$$
$$\pi_i = \text{logit}^{-1}(\beta_0+\beta_1 x_{i}),$$
$$\beta_0 \sim N(0,100), \;  \beta_1 \sim N(0,100),$$
where $y_i, x_{i}$ denote the nodes detected of the person $i$.


#### Put your code below

Prepare the data 

```{r}
biopsy_dat <- list(N=length(biopsy$age), y=biopsy$survival, x=biopsy$nodes)
```

Hint: Note that the observations are now binary. So in the data section y should be declared as

int<lower=0,upper=1> y[N];

x should also be declared in the same way as y in the toy examples

vector[N] x;